A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2019-06-10 paper/paper.Rnw
Update outline and keep reorganizing and filling in results

\mt{The model base of VGG16 is made up of five convolutional blocks alternated with max pooling layers. Each convolutional block is made up of two or three convolutional layers, and each layer is made up of hundreds of filters, such that each of the five convolutional blocks has between 128 and 1,536 filters. Strictly speaking, a filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect. \autoref{fig:vgg16layers} shows a selection of filters from VGG16, with each row corresponding to one of the five convolutional blocks. Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features. When considered together, the series of filters perform the process of feature detection, such that a single image is made up of some combination of the thousands of features that are detected by filters. Although the filters of VGG16 have been trained and optimized on the ImageNet Data, the large number of filters span a wide variety of shapes and features} \fix{which are still useful/relevant to classifying outsoles. In a sense, this is much like teaching humans to classify objects they have never seen before: an object may be unfamiliar, but our eyes can still detect its color, shape, and texture because we have learned these ideas from other sources.}

@@

%\svp{Moving towards ... how did we train the model}
\paragraph{Annotated Training Data}

@@

%Labeled images are scarce relative to the amount of data necessary to train a neural network.
Although transfer learning substantially reduces the amount of data required to train a CNN by several orders of magnitude, labeled images are still \fix{difficult or time-consuming} to generate relative to the amount of data necessary to train even a small neural network adequately. Thus, to make the most efficient use of the existing labeled data, we enlarge the training data using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}. Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label. This process reduces the potential for overfitting the model to the specific set of image data used during the training process, and also increases the amount of data available for training. Examples of pre- and post-augmentation images are shown in \autoref{fig:augmented}.

@@

\mt{\textbf{OUTLINE OF RESULTS}
\begin{itemize}
\item ROC
  \begin{itemize}
  \item Overall: model performance is better than chance
  \item Classes: broad comparison of class performance
  \end{itemize}
\item Examples
  \begin{itemize}
  \item Show what predictions look like
  \item Introduce the idea that some classifications aren't perfect
  \item Motivate systematic class-by-class confusion
  \end{itemize}
\item Confusion Matrix
  \begin{itemize}
  \item Show patterns in misclassification
  \item Conf mat doesn't tell us why certain predictions are wrong
  \end{itemize}
\item In this paper, we identify two main reasons for error: some classes are more variable, and some classes share physical features (I.e., the model has to learn a lot of features to encapsulate a class with lots of variety, but then other classes share those features)
\item Brands
  \begin{itemize}
  \item Bowties/polygons/stars are mostly a few brands, but the top quad/line brands still only take up a small proportion of all quads (.04 or less)
  \end{itemize}
\item Heatmaps
  \begin{itemize}
  \item Verify that the model is using appropriate features when predictions are correct
  \item Show examples for strong misclassification, but make the case that the mistakes are understandable
  \end{itemize}
\end{itemize}
}

@@

For each image in the test set, the ``true labels" are nine human-assigned labels of 0 or 1 corresponding to the presence or absence of each of the nine shape categories in the image. When the CNN predicts which shapes are in the image, however, it assigns a probability between 0 and 1 for each shape category, with a total of nine probabilities per test image. Thus, to determine the accuracy of the model predictions, we must select a threshold for each label to discretize the model predictions.
If the threshold is too high, \fix{moderate} probabilities will be treated as a ``no", which increases the number of true negative and false negative predictions. Conversely, if the threshold is too low, \fix{moderate} probabilites will be considered as ``yes", which increases the true and false positive predictions in the data. An appropriate threshold must be chosen to produce a high true positive rate while also ideally keeping the false positive rate low.

@@

\paragraph{Model Predictions}
\mt{ROC curves provide a general summary of model accuracy. To learn more about how specific images are classified, we can view images from the test set along with their respective nine model predictions. \autoref{fig:single-exemplars} shows model predictions for images that contain only one type of shape, and \autoref{fig:multi-exemplars} shows similar predictions for images containing multiple shapes. Most predictions for the single-label images are straightforward, but there is much less consensus of model opinion in the multi-label images. }

@@

@@

\mt{These figures provide representative examples of classifications across multiple categories; it is also useful to examine many cases within a given shape. \autoref{fig:chevrons-consistency}}

@@

@@

\fix{Now that we've seen specific examples, it is useful to summarize the systematic errors by class.} A confusion matrix is a helpful diagnostic tool to understand which classes are commonly confused for other classes.} A confusion matrix is a \mt{summary of the relationships between} the labels assigned to an image and the labels predicted by the model.

In the confusion matrix presented in \autoref{fig:confusion_matrix}, the values along the diagonal represent the proportion of true positives captured within each category. The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values have been adjusted to remove the effect of any true positives from the calculation of false positive proportions. \svp{As a result of this adjustment, the probabilities off-diagonal are the proportion ... need to explain the conditional probability we're showing here} %\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected}
For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.

@@

\autoref{fig:confusion_matrix} shows systematic errors in class predictions, but does not necessarily suggest \emph{why} the confusion is taking place. To assess the reasons for model failure, we look at the images to \fix{see what the model is seeing.}

\svp{I would move heatmaps up and out of the diagnostic section and transition into that, possibly}

\svp{Three main model exploration chunks: consistency, heatmaps, brand (motivated by the fact that some classes perform so badly because the exemplars are so variable) - need to figure out the best sequence to transition}

@@

A heatmap \fix{please add the technical name for heatmaps - look in the distill.io papers?} is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class. Using these graphics, we can identify which locations in an image contribute most to the model's predictions. In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.

>>>>

\mt{The model base of VGG16 is made up of five convolutional blocks alternated with max pooling layers. Each convolutional block is made up of two or three convolutional layers, and each layer is made up of hundreds of filters, such that each of the five convolutional blocks has between 128 and 1,536 filters. Strictly speaking, a filter is a set of numerical weights that, when applied to an image, quantifies the presence of the specific shapes and/or colors that the filter is trained to detect. \fix{Necessary? (Or just skip...): Visualizing a filter then refers to visualizing the features to which the filter provides its maximum response.} \autoref{fig:vgg16layers} shows a selection of filters from VGG16, with each row corresponding to one of the five convolutional blocks. Filters in earlier convolutional blocks detect simple features, such as colors and blobs, while later filters detect more complex combinations of features. When considered together, the series of filters perform the process of feature detection, such that a single image is made up of some combination of the thousands of features that are detected by filters. Although the filters of VGG16 have been trained and optimized on the ImageNet Data, the large number of filters span a wide variety of shapes and features} \fix{which are still useful/relevant to classifying outsoles. In a sense, this is much like teaching humans to classify objects they have never seen before: an object may be unfamiliar, but our eyes can still detect its color, shape, and texture because we have learned these ideas from other sources.}

@@

\paragraph{Annotated Training Data}

@@

Although transfer learning substantially reduces the amount of data required to train a CNN by several orders of magnitude, labeled images are still \fix{difficult or time-consuming} to generate relative to the amount of data necessary to train even a small neural network adequately. Thus, to make the most efficient use of the existing labeled data, we enlarge the training data using a process called image augmentation \citep{krizhevskyImageNetClassificationDeep2012}. Augmentation is the transformation of original input data using image operations such as cropping, zoom, skew, rotation, and color balance modification in order to distort or alter the image while maintaining the essential features corresponding to the label. This process reduces the potential for overfitting the model to the specific set of image data used during the training process, and also increases the amount of data available for training. Examples of pre- and post-augmentation images are shown in \autoref{fig:augmented}.

@@

\mt{\textbf{OUTLINE OF RESULTS}
\begin{itemize}
\item ROC
  \begin{itemize}
  \item Overall: model performance is better than chance
  \item Classes: broad comparison of class performance
  \end{itemize}
\item Examples
  \begin{itemize}
  \item Show what predictions look like
  \item Introduce the idea that some classifications aren't perfect
  \item Motivate systematic class-by-class confusion
  \end{itemize}
\item Confusion Matrix
  \begin{itemize}
  \item Show patterns in misclassification
  \item Motivates further understanding of common mistakes
  \end{itemize}
\item Two main reasons for error: some classes are more variable, and some classes share physical features (I.e., the model has to learn a lot of features to encapsulate a class with lots of variety, but then other classes share those features)
\item Brands
  \begin{itemize}
  \item Bowties/polygons/stars are mostly found in only a few brands, but the top N quad/line brands only take up a small proportion of all quads/lines (.04 or less)
  \item In general, the most difficult classes to classify are those that are highly variable
  \end{itemize}
\item Heatmaps
  \begin{itemize}
  \item Verify that the model is using appropriate features when predictions are correct
  \item Show examples for confident misclassifications (like Seychelles), but make the case that the mistakes are understandable
  \end{itemize}
\end{itemize}
}

@@

For each image in the test set, the ``true" labels are nine human-assigned labels of 0 or 1 corresponding to the presence or absence of each of the nine shape categories in the image. When the CNN predicts which shapes are in the image, however, it assigns a probability between 0 and 1 for each shape category, with a total of nine probabilities per test image. Thus, to determine the accuracy of the model predictions, we must select a threshold for each label to discretize the model predictions.
If the threshold is too high, \fix{moderate} probabilities will be treated as a ``no", which increases the number of true negative and false negative predictions. Conversely, if the threshold is too low, \fix{moderate} probabilites will be considered as ``yes", which increases the true and false positive predictions in the data. An appropriate threshold must be chosen to produce a high true positive rate while also ideally keeping the false positive rate low.

@@

\paragraph{Model Predictions}
\mt{The ROC curves provide only a general summary of model accuracy; to learn more about how specific images are classified, we can view images from the test set along with their respective nine model predictions.} \autoref{fig:single-exemplars} shows model predictions for images that contain only one type of shape, and \autoref{fig:multi-exemplars} shows similar predictions for images containing multiple shapes. Most predictions for the single-label images are straightforward; all of the archetypical shapes are corectly identified by the model, with only small confusions like those between quad/polygon and circle/text. \fix{A later section will discuss possible causes for this confusion.}

@@

The multi-label images also perform quite well. \fix{Find a few more examples here and discuss them. Maybe a few similar ones to show consistency matters more than explicit label. Make the point that this task is messy and shape definitions are hard, but the model is doing pretty well given the circumstances.}

@@

\fix{These figures provide representative examples of classifications across multiple categories; it is also useful to examine many cases within a given shape.} \mt{\autoref{fig:chevrons-consistency} shows a number of images containing chevrons. The top four rows show regularly repeating chevron patterns and seems to indicate that the model optimally identifies chevrons of a certain width; wide chevrons cross into the realm of bowties, whereas narrow chevrons become indistinguishable from lines. The next six rows show individual and closed chevrons. There is a significant amount of confusion between these wider chevrons and quadrilaterals. Finally, the bottom four rows show less-than-standard chevrons. The rounded chevron has a much lower prediction than those with sharper angles, but still exceeds the EER for chevron; the white chevrons are confounded with overlaid circles, which introduces a complexity that the model attributes to text.}

@@

\fix{Describe \autoref{fig:circle-text-consistency}...... The final two don't contain a true circle. IDK why the TOMS logo totally misses circle.}

@@

mt{\paragraph{Accuracy by class}}
\fix{Many of the trends in the examples in figures \ref{fig:single-exemplars}-\ref{fig:circle-text-consistency} are not unique. A confusion matrix is a type of visualization that summarizes the relationships between the labels assigned to an image and the labels predicted by the model, and is a helpful diagnostic tool to understand which classes are systematically misclassified.} \mt{The confusion matrix for \fix{our model} is presented in \autoref{fig:confusion_matrix}}. The values along the diagonal represent the proportion of true positives captured within each category. The off-diagonal values represent the proportion of false positives for each shape pairing; however, since a single image may truly contain multiple shapes, the values have been adjusted to remove the effect of any true positives from the calculation of false positive proportions.
%\svp{As a result of this adjustment, the probabilities off-diagonal are the proportion ... need to explain the conditional probability we're showing here}
\fix{As a result of this adjustment, the off-diagonal values are the proportion of images within a true class that are falsely labeled as containing another class that has not been identified in the image.}%\svp{I can't remember exactly how we're doing that, but we may want to look more closely at that - if we set everything to NA instead of 0, the rate calculations would be higher than expected}
For example, to calculate the proportion of images that contain triangles but are being falsely labeled as containing quadrilaterals, any image that truly contains both triangles and quadrilaterals is removed before calculating the proportion of false quadrilateral labels.

@@

%\autoref{fig:confusion_matrix} shows systematic errors in class predictions, but does not necessarily suggest \emph{why} the confusion is taking place. To assess the reasons for model failure, we look at the images to \fix{see what the model is seeing.}

%\svp{I would move heatmaps up and out of the diagnostic section and transition into that, possibly}

%\svp{Three main model exploration chunks: consistency, heatmaps, brand (motivated by the fact that some classes perform so badly because the exemplars are so variable) - need to figure out the best sequence to transition}

@@

Belonging to a category of techniques called class activation map (CAM) visualizations, a heatmap is a visual diagnostic tool that highlights the portions of an image that are most significant to classification for a given class \citep{chollet_allaire_2018}. Using these graphics, we can identify which locations in an image contribute most to the model's predictions. In \fix{the figures in this subsection,} heatmaps are shown for a selection of three classes, which are either contained in the image or have high output probabilities, alongside the original image.
